---
title: "R code for 'Asymmetric growth responses to changes in spring and autumn phenology in temperate woody plants'"
---

<br><br>

## 1. Load packages

<details><summary>get packages</summary>

```{r, message=FALSE, warning=FALSE}
require(ggplot2)
require(dplyr)
require(patchwork)
require(lubridate)
require(data.table)
require(git2r)
require(stringr)
require(broom)
require(viridis)
require(car)
```
</details>

<br>

## 2. Load data	

<details><summary>data tables</summary>

```{r}
# define paths
data.dir   = "/Users/consti/Desktop/PhD/Publication_material/10_PhenogrowthEx3/Rdata/Experiment"
output.dir = "/Users/consti/Desktop/PhD/Publication_material/10_PhenogrowthEx3/Routput"

#load data
clim.data    = data.frame(fread(paste(data.dir,"climate_data.csv",sep="/"), header=T, sep=",")) # Climate 
volume.data  = data.frame(fread(paste(data.dir,"volume.csv",sep="/"), header=T, sep=","))       # Stem volume
pheno.data   = data.frame(fread(paste(data.dir,"phenology.csv",sep="/")))                       # Phenology
biomass.data = data.frame(fread(paste(data.dir,"biomass.csv",sep="/")))                         # Biomass
```
</details>

<details><summary>ggplot themes</summary>
```{r}
#define plot themes
plotTheme = theme(legend.position   = "right",
                  legend.background = element_rect(fill=NA, size=0.5, linetype="solid"),
                  legend.text       = element_text(color="black"),
                  panel.grid.major  = element_blank(),
                  panel.grid.minor  = element_blank(),
                  panel.background  = element_rect(colour = NA, size=1, fill="grey97"),
                  axis.line         = element_line(color = "black"),
                  strip.background  = element_rect(fill=NA),
                  strip.text        = element_text(colour = 'black'))

plotTheme2 = theme(legend.position   = "none",
                   legend.background = element_rect(fill=NA, size=0.5, linetype="solid"),
                   legend.text       = element_text(color="black"),
                   panel.grid.major  = element_blank(),
                   panel.grid.minor  = element_blank(),
                   panel.background  = element_rect(colour = NA, size=1, fill="grey97"),
                   axis.text         = element_blank(),
                   axis.ticks        = element_blank(),
                   strip.background  = element_rect(fill=NA),
                   strip.text        = element_text(colour = 'black'))

plotTheme3 = theme(legend.position   = "none",
                  legend.background = element_rect(fill=NA, size=0.5, linetype="solid"),
                  legend.text       = element_text(color="black"),
                  panel.grid.major  = element_blank(),
                  panel.grid.minor  = element_blank(),
                  panel.background  = element_rect(colour = NA, size=1, fill="grey97"),
                  axis.line         = element_line(color = "black"),
                  strip.background  = element_rect(fill=NA),
                  strip.text        = element_text(colour = 'black'))

plotTheme4 = theme(legend.position   = "right",
                  legend.background = element_rect(fill=NA, size=0.5, linetype="solid"),
                  legend.text       = element_text(color="black"),
                  panel.grid.major  = element_blank(),
                  panel.grid.minor  = element_blank(),
                  panel.background  = element_rect(colour = NA, size=1, fill="grey97"),
                  axis.text.x       = element_text(angle = 45, hjust = 1),
                  axis.line.y       = element_line(color = "black"),
                  strip.background  = element_rect(fill=NA),
                  strip.text        = element_text(colour = 'black'))
```
</details>

<br>

## 3. Data preparation	

<details><summary>climate data</summary>
<details><summary>data management</summary>
```{r, message=FALSE, warning=FALSE}
#Convert dates to DOY
clim.data$date = dmy(clim.data$date)

#get median deviations from control + interquartile ranges
clim.data1 = data.frame(clim.data %>%
                     filter(!is.na(inside_tmp)) %>% #delete summer period
                     mutate(diffOutIn  = inside_tmp-outside_tmp)) %>% #add deviation column
  group_by(treatment) %>%
  summarize(medianDiff = median(diffOutIn),
            IQrange    = IQR(diffOutIn, na.rm = FALSE, type = 7))

#get new dataframe for deviation plot
clim.data2 = data.frame(clim.data %>%
                     filter(!is.na(inside_tmp)) %>% #delete NAs
                     mutate(diffOutIn  = inside_tmp-outside_tmp)) #get deviation from control

#reshape initial dataframe
clim.data3 = melt(clim.data, id.vars = c("date","treatment"), measure.vars = c("outside_tmp","inside_tmp"))

#create treatment column
clim.data3$variable2 = ifelse(clim.data3$variable=="outside_tmp", "control", 
                       ifelse(clim.data$treatment=="autumn", "autumn", "spring"))
```
</details>

<details><summary>temperature table</summary>
```{r, message=FALSE, warning=FALSE}
data.frame(clim.data1)
```
</details>
</details>



<details><summary>stem volume</summary>
<details><summary>stem volume calculation</summary>
```{r, warning=FALSE}
#order treatments
Treatment = c("spring", "autumn", "autumn-spring", "control")
volume.data$Treatment = factor(volume.data$Treatment, levels=Treatment)

#calculate volumes
volume.data$Volume = 1/3 * pi * volume.data$Height *
  ((volume.data$StemDiameter/2)^2 + 
     (volume.data$StemDiameter/2)*(volume.data$StemDiameterTop/2) + 
     volume.data$StemDiameterTop^2)
                        
#sum stem and twig volume measurements within individuals for each year
VariablesVector = c("Volume")
volume.data = data.frame(volume.data %>% 
    group_by(Species,ID,Treatment,Year) %>% 
    summarize_at(VariablesVector, sum, na.rm = TRUE))
```
</details>
<details><summary>sample sizes</summary>
```{r, message=FALSE, warning=FALSE}
table(volume.data$Treatment, volume.data$Species)/4
```
</details>
</details>



<details><summary>phenology</summary>
```{r, warning=FALSE, message=FALSE}
#get RMFs
biomass.data$RSR = biomass.data$Root/biomass.data$Shoot

#delete species epithet in all tables
DeleteEpithet <- function(df) {
    df$Species = gsub(' [A-z ]*', '' , df$Species)
    return(df)
}
dfnames    = list(pheno.data,biomass.data,volume.data)
dfs        = lapply(dfnames, DeleteEpithet)
names(dfs) = c("pheno.data","biomass.data","volume.data")

#Ordering of factors
Treatments      = c("control", "autumn", "autumn-spring", "spring") #order treatments
pheno.data$Treatment = factor(pheno.data$Treatment, levels=Treatments, ordered=T)
Species         = c("Quercus", "Fagus", "Lonicera") #order species
pheno.data$Species   = factor(pheno.data$Species, levels=Species, ordered=T)

#reshape phenology table to long format
pheno.data       = melt(dfs$pheno.data, id.vars = c("Species","ID","Treatment"))

#Convert dates to DOY
pheno.data$value = dmy(pheno.data$value)
pheno.data$DOY   = yday(pheno.data$value) # Jan 1 = day 1 

#create year and phenophase columns
pheno.data$Year  = str_sub(pheno.data$variable, str_length(pheno.data$variable)-3, str_length(pheno.data$variable))
pheno.data$Pheno = str_sub(pheno.data$variable, 1, str_length(pheno.data$variable)-5)
pheno.data       = select(pheno.data, -c(variable, value))
pheno.data       = na.omit(pheno.data)

#from long (many rows) to short (multiple columns) format
pheno.data       = dcast(pheno.data, Species + ID + Treatment + Year ~ Pheno, value.var = "DOY")

#get growing season length information (leaf-out to senescence)
pheno.data$GSL   = pheno.data$Senescence - pheno.data$Leaf_out

#constrain late leafers
pheno.data$Leaf_out = ifelse(pheno.data$Leaf_out>140,140,pheno.data$Leaf_out)
```
</details>



<details><summary>biomass</summary>
```{r, warning=FALSE, message=FALSE}
#Melt data to a long format
biomass.data       = melt(dfs$biomass.data, id = c("Species", "ID", "Treatment"), 
                  variable.name="organ", value.name="biomass")

#Ordering of factors
Treatments     = c("control", "autumn", "autumn-spring", "spring") #order treatments
biomass.data$Treatment = factor(biomass.data$Treatment, levels=Treatments, ordered=T)
Species        = c("Quercus", "Fagus", "Lonicera") #order species
biomass.data$Species   = factor(biomass.data$Species, levels=Species, ordered=T)
variable       = c("Total", "Shoot", "Root", "RSR") #order organ
biomass.data$organ     = factor(biomass.data$organ, levels=variable, ordered=T)
```
</details>

<br>

## 4. Analysis

<details><summary>phenology</summary>
<details><summary>treatment effect on phenology</summary>
```{r, warning=FALSE, message=FALSE}
#################################
# Summarize annual phenology data
#################################


#get average phenology anomalies per individual
###############################################

#reshape phenology columns to long format
pheno.data = melt(pheno.data, id.vars = c("Species","ID","Treatment","Year"), 
                  variable.name="phenology", value.name="day")

#get means per species and year
phenoMean = pheno.data %>% 
    group_by(Species,phenology,Year) %>% 
    summarize(Mean = mean(day, na.rm=TRUE))

#merge data
pheno.data <-merge(pheno.data, phenoMean, all.x=T, by=c("Species","phenology","Year"))
rm(phenoMean)

#get change in phenology relative to mean
pheno.data = pheno.data %>% 
  mutate(phenologyAnomaly = ifelse(phenology=="Leaf_out", Mean-day, day-Mean)) %>% #add column
  dplyr::select(!c(Mean)) #delete columns

#from long (many rows) to short (multiple columns) format
pheno.data3 = dcast(pheno.data, Species + ID + Treatment + Year ~ phenology, value.var = c("phenologyAnomaly"))

#summarize years
VariablesVector = c("Leaf_out","Senescence","GSL")
pheno.data3 = data.frame(pheno.data3 %>%
    filter(!Year %in% c("2017")) %>% #delete rows based on condition
    group_by(Species,ID,Treatment) %>% 
    summarize_at(VariablesVector, mean, na.rm = TRUE))


# Add biomass information
#########################

#merge phenology and biomass tables
pheno.data3 = merge(pheno.data3, dfs$biomass.data, all.x=T, by=c("Species","ID","Treatment"))

#reshape phenology columns to long format
pheno.data3 = melt(pheno.data3, id.vars = c("Species","ID","Treatment","Total","Root","Shoot","RSR"), 
             variable.name="phenology", value.name="phenologyAnomaly")


#get average phenology dates per individual
###########################################

#from long (many rows) to short (multiple columns) format
pheno.data1 = dcast(pheno.data, Species + ID + Treatment + Year ~ phenology, value.var = c("day"))

#summarize years
VariablesVector = c("Leaf_out","Senescence","GSL")
pheno.data1 = data.frame(pheno.data1 %>%
    filter(!Year %in% c("2017")) %>% #delete rows based on condition
    group_by(Species,ID,Treatment) %>% 
    summarize_at(VariablesVector, mean, na.rm = TRUE))

#merge phenology and biomass tables
pheno.data1 = merge(pheno.data1, dfs$biomass, all.x=T, by=c("Species","ID","Treatment"))

#reshape phenology columns to long format
pheno.data1 = melt(pheno.data1, id.vars = c("Species","ID","Treatment","Total","Root","Shoot","RSR"), 
             variable.name="phenology", value.name="day")


#########################################################################################
#########################################################################################


###########################
# Multivariate linear model
###########################


#Create spring and autumn treatments
####################################

pheno.data4        = pheno.data3
pheno.data4$spring = ifelse(pheno.data4$Treatment %in% c("control","autumn"), "no","warming")
pheno.data4$autumn = ifelse(pheno.data4$Treatment %in% c("control","spring"), "no","warming")

#Multivariate linear model
##########################

resultsLM.pheno = pheno.data4 %>% 
    group_by(Species,phenology) %>% 
    do({model = lm(phenologyAnomaly ~ spring*autumn, data=.)  # create your model
    data.frame(tidy(model))}) %>%           # get coefficient info
    filter(!term %in% c("(Intercept)")) %>% # delete rows
    dplyr::select(Species, phenology, term, p.value, estimate, std.error) %>% #delete columns
    mutate_if(is.numeric, round, 2) %>%   
    mutate(sig = ifelse(p.value<0.05, "**", ifelse(p.value<=0.1, "*", ""))) #add significance
#Ordering of factors
term                 = c("springwarming", "autumnwarming", "springwarming:autumnwarming") #order treatments
resultsLM.pheno$term = factor(resultsLM.pheno$term, levels=term, ordered=T)
Species                 = c("Quercus", "Fagus", "Lonicera") #order species
resultsLM.pheno$Species = factor(resultsLM.pheno$Species, levels=Species, ordered=T)
#order statistics table to match dataframe
resultsLM.pheno = resultsLM.pheno[with(resultsLM.pheno, order(resultsLM.pheno$Species, resultsLM.pheno$phenology)), ]
data.frame(resultsLM.pheno[,-c(7)])



######################################################
# Check for deviation from control (one-sample T-test)
######################################################

#get means of control group
pheno.data2 = pheno.data1 %>% 
    filter(Treatment %in% c("control")) %>%
    group_by(Species,phenology) %>% 
    summarize(Mean = mean(day, na.rm=TRUE))
#merge data
pheno.data2 <-merge(pheno.data1[!pheno.data1$Treatment=="control",], 
                    pheno.data2, all.x=T, by=c("Species","phenology"))
#get change in phenology relative to control
pheno.data2$phenologyChange = pheno.data2$day - pheno.data2$Mean

#Ordering of factors
Phenology     = c("Leaf_out", "Senescence", "GSL") #order treatments
pheno.data2$phenology = factor(pheno.data2$phenology, levels=Phenology, ordered=T)
Treatments     = c("spring", "autumn", "autumn-spring") #order treatments
pheno.data2$Treatment = factor(pheno.data2$Treatment, levels=Treatments, ordered=T)
Species         = c("Quercus", "Fagus", "Lonicera") #order species
pheno.data2$Species   = factor(pheno.data2$Species, levels=Species, ordered=T)

#T-test
resultsTT = pheno.data2 %>% 
    group_by(Species,phenology,Treatment) %>% 
    do({model = t.test(.$phenologyChange) # create your model
    data.frame(tidy(model),
               tidy(shapiro.test(x = .$phenologyChange))[1,2])}) %>%           
    dplyr::select(Species, phenology, Treatment, p.value, estimate, p.value.1) %>% #delete columns
    rename(shapiroTest=p.value.1) %>%      
    mutate(significance = ifelse(p.value<0.05, "**", ifelse(p.value<=0.1, "*", ""))) %>% #add significance
    mutate_if(is.numeric, round, 2) 
#order statistics table to match dataframe
resultsTT = resultsTT[with(resultsTT, order(resultsTT$Species, resultsTT$phenology)), ]
data.frame(resultsTT[,-c(7)])
```
</details>



<details><summary>phenology effect on total biomass</summary>
```{r, warning=FALSE, message=FALSE}
##########################
# Univariate linear models
##########################

#reshape biomass columns to long format
pheno.data3 = melt(pheno.data3, id.vars = c("Species","ID","Treatment","phenology","phenologyAnomaly"), 
             variable.name="organ", value.name="biomass")

#Transform biomass to percentage anomaly
########################################

#get means per species
pheno.data5 = pheno.data3 %>% 
    group_by(Species,organ) %>% 
    summarize(Mean = mean(biomass, na.rm=TRUE))
#merge data
pheno.data3 <-merge(pheno.data3, pheno.data5, all.x=T, by=c("Species","organ"))
#get precentage change in biomass
pheno.data3$relativeBiomass = (pheno.data3$biomass / pheno.data3$Mean -1) * 100

#order species
Species             = c("Quercus", "Fagus", "Lonicera") 
pheno.data3$Species = factor(pheno.data3$Species, levels=Species, ordered=T)

#Extract linear model info
resultsLM.pheno2 = pheno.data3 %>% 
    group_by(Species,phenology,organ) %>% 
    do({model = lm(relativeBiomass ~ phenologyAnomaly, data=.)  # create model
    data.frame(tidy(model),                # get coefficient info
               glance(model),
               tidy(shapiro.test(x = residuals(object = model)))[1,2]
               )}) %>%                     # get model info
    filter(term != "(Intercept)") %>%      # delete intercept info
    dplyr::select(Species, phenology, organ, estimate, std.error, p.value, r.squared, p.value.2) %>% # delete columns
    rename(shapiro=p.value.2) %>%          #rename columns
    mutate(sig = ifelse(p.value<0.05, "**", ifelse(p.value<=0.1, "*", ""))) %>% # add column
    mutate_if(is.numeric, round, 2)

#Order phenophases
Organs            = c("Total","Shoot","Root","RSR") #order treatments
resultsLM.pheno2$organ   = factor(resultsLM.pheno2$organ, levels=Organs, ordered=T)
#order statistics table to match dataframe
resultsLM.pheno2 = resultsLM.pheno2[with(resultsLM.pheno2, 
                                         order(resultsLM.pheno2$Species, 
                                               resultsLM.pheno2$organ, 
                                               resultsLM.pheno2$phenology)), ]
data.frame(resultsLM.pheno2)

#Visually inspect model assumptions 
data.assumptions = pheno.data3
data.assumptions$category = paste(data.assumptions$Species,
                                  data.assumptions$organ,
                                  data.assumptions$phenology, sep="_") #create identifier column
category.list = as.factor(unique(data.assumptions$category))       #create category vector
par(mfrow=c(2,2))                                       #set plot layout
for (category in category.list){                        #loop over categories
  tab.subset=data.assumptions[data.assumptions$category==category, ] #create table subset
  plot(lm(relativeBiomass ~ phenologyAnomaly, data=tab.subset), main=category)
}
```
</details>
</details>



<details><summary>biomass</summary>
<details><summary>total and per treatment biomass distributions</summary>
```{r}
#Species-level plot
ggplot(data = biomass.data, mapping = aes(biomass, fill=organ)) +
  geom_histogram(bins=10, position="identity", color="black") +
  xlab("Biomass (g) [panels 1-3]                     
                                                                                           Biomass ratio [panel 4]") +
  ylab("Count") +
  scale_fill_viridis(option="E",discrete=TRUE) +
  facet_grid(Species~organ, scales="free") + 
  plotTheme4

#Treatment-level plot
ggplot(data = biomass.data[biomass.data$organ %in% c("Shoot","Root"),], mapping = aes(biomass, fill=organ)) +
  geom_histogram(bins=10, color="black") +
  xlab("Biomass (g)") +
  ylab("Count") +
  scale_fill_viridis(option="C",discrete=TRUE) +
  facet_grid(organ+Species~Treatment, scales="free") + 
  plotTheme4
```
</details>



<details><summary>treatment analysis</summary>
```{r}
#######
# ANOVA
#######

#Extract model info and check ANOVA assumptions
resultsLM = biomass.data %>% 
    group_by(Species,organ) %>% 
    do({model = aov(biomass ~ Treatment, data=.)        # create your model
    data.frame(tidy(model)[1,],                         # get coefficient info
               glance(model)[1,],                       # get model info
               leveneTest=tidy(leveneTest(model))[1,4], # check Homogeneity of variances
               tidy(shapiro.test(x = residuals(object = model)))[1,2] # check for Normality
               )}) %>%           
    #select(Species, organ, p.value, r.squared, adj.r.squared, p.value.2, p.value.3) %>% #delete columns
    rename(leveneTest=p.value.2, shapiroTest=p.value.3) %>%                             #rename columns
    mutate(sig = ifelse(p.value<0.05, "**", ifelse(p.value<0.1, "*", "")),
           Treatment="control")     #add significance and dummy column for plotting
data.frame(resultsLM)

# Plots to check ANOVA assumptions
biomass.data$category = paste(biomass.data$Species,biomass.data$organ, sep="_") #create identifier column
category.list = as.factor(unique(biomass.data$category))        #create category vector
par(mfrow=c(2,2))                                       #set plot layout
for (category in category.list){                        #loop over categories
  tab.subset=biomass.data[biomass.data$category==category, ]            #create table subset
  plot(aov(biomass ~ Treatment, data=tab.subset), 1, main=category) # 1. Homogeneity of variances
  plot(aov(biomass ~ Treatment, data=tab.subset), 2, main=category) # 2. Normality
}


#####
#Plot
#####

Boxp <- ggplot(data = biomass.data, mapping = aes(x = Treatment, y = biomass, fill=Treatment)) +
  geom_boxplot(position=position_dodge(0.8),coef=1e30) +
  geom_text(data = resultsLM,
            mapping = aes(x = Inf, y = Inf, hjust = 2, vjust = 2, 
                            label = paste("R2 = ", round(r.squared,2), sig, sep="")), 
              size=3.5, color="black")+
  xlab("Treatment") +
  ylab("Biomass (g)") +
  scale_fill_viridis(option="E",discrete=TRUE) +
  facet_grid(organ~Species, scales="free") + 
  plotTheme4
Boxp
```
</details>



<details><summary>Relative biomass analysis</summary>
```{r}
###########################
# Multivariate linear model
###########################


#Create spring and autumn treatments
####################################

biomass.data$spring = ifelse(biomass.data$Treatment %in% c("control","autumn"), "no","warming")
biomass.data$autumn = ifelse(biomass.data$Treatment %in% c("control","spring"), "no","warming")


#Transform biomass to percentage anomaly
########################################

#get means per species
biomass.data1 = biomass.data %>% 
                group_by(Species,organ) %>% 
                summarize(Mean = mean(biomass, na.rm=TRUE))
#merge data
biomass.data1 <-merge(biomass.data, biomass.data1, all.x=T, by=c("Species","organ"))
#get precentage change in biomass
biomass.data1$relativeBiomass = (biomass.data1$biomass / biomass.data1$Mean -1) * 100


#Multivariate linear model
##########################

resultsLM.biomass = biomass.data1 %>% 
    group_by(Species,organ) %>% 
    do({model = lm(relativeBiomass ~ spring*autumn, data=.)  # create your model
    data.frame(tidy(model))}) %>%           # get coefficient info
    filter(!term %in% c("(Intercept)")) %>% # delete rows
    dplyr::select(Species, organ, term, p.value, estimate, std.error) %>% #delete columns
    mutate_if(is.numeric, round, 2) %>%   
    mutate(sig = ifelse(p.value<0.05, "**", ifelse(p.value<=0.1, "*", ""))) #add significance
data.frame(resultsLM.biomass)


########
# T-test
########

#get means of control group
###########################

biomass.data1 = biomass.data %>% 
    filter(Treatment %in% c("control")) %>%
    group_by(Species,organ) %>% 
    summarize(Mean = mean(biomass, na.rm=TRUE))
#merge data
biomass.data1 <-merge(biomass.data[!biomass.data$Treatment=="control",], biomass.data1, all.x=T, by=c("Species","organ"))
#get precentage change in biomass
biomass.data1$relativeBiomass = ifelse(biomass.data1$organ!="RSR", (biomass.data1$biomass / biomass.data1$Mean -1) * 100,
                                      (biomass.data1$biomass - biomass.data1$Mean)*100)
#Ordering of factors
Treatments     = c("spring", "autumn", "autumn-spring") #order treatments
biomass.data1$Treatment = factor(biomass.data1$Treatment, levels=Treatments, ordered=T)
Species     = c("Quercus", "Fagus", "Lonicera") #order treatments
biomass.data1$Species = factor(biomass.data1$Species, levels=Species, ordered=T)

#T-test
#######

resultsTTbiomass = biomass.data1 %>% 
    group_by(Species,organ,Treatment) %>% 
    do({model = t.test(.$relativeBiomass)        # create your model
    data.frame(tidy(model),
               tidy(shapiro.test(x = .$relativeBiomass))[1,2])}) %>%           
    select(Species, organ, Treatment, p.value, estimate, p.value.1) %>% #delete columns
    rename(shapiroTest=p.value.1) %>%   
    mutate_if(is.numeric, round, 2) %>%   
    mutate(sig = ifelse(p.value<0.05, "**", ifelse(p.value<=0.1, "*", ""))) #add significance
#order statistics table to match dataframe
resultsTTbiomass = resultsTTbiomass[with(resultsTTbiomass, 
                                         order(resultsTTbiomass$Species, resultsTTbiomass$organ)), ]
data.frame(resultsTTbiomass)
```
</details>
</details>

<br>

## 5. Figures

<details><summary>Figure 1</summary>
```{r, message=FALSE, warning=FALSE}
#2D density plot
Dens2D = ggplot(clim.data, aes(x=outside_tmp, y=inside_tmp-outside_tmp) ) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon") +
  scale_fill_distiller(palette=4, direction=1) +
  geom_hline(aes(yintercept=4.7))+
  coord_cartesian(ylim = c(0, 10),xlim = c(-4, 19))+
   xlab("Ambient temperature") +
  ylab("Deviation from control (C)") +
  plotTheme

#Boxplot
boxp = ggplot(clim.data2, aes(x=treatment, y=diffOutIn, fill=treatment)) + 
  geom_boxplot(outlier.shape=NA,alpha=.6) +
  coord_flip(ylim = c(0.5, 10)) +
  xlab("") +
  ylab("") +
  scale_fill_manual(values=c("orange","green3"))+
  plotTheme2

#Histogram
histo = ggplot(clim.data2, aes(x=diffOutIn, fill=treatment)) +
  geom_histogram(binwidth=.5, alpha=.6, position="identity") +
  geom_vline(data=clim.data1, aes(xintercept=medianDiff, colour=c("green3","orange"),alpha=.8),
             linetype="dashed", size=1)+
  xlab("Deviation from control (C)") +
  ylab("Count (number of days)") +
  scale_fill_manual(values=c("orange","green3"))+
    scale_colour_manual(values=c("orange","green3"))+
  coord_cartesian(xlim = c(0.5, 10), ylim = c(3, 60))+
  plotTheme3

#Time plot
TempPlot = ggplot(clim.data3, aes(x = date, y= value, group = variable, colour = variable2)) +
  geom_line() +
  scale_color_manual(values=c("orange","black","green3"))+
  xlab("") +
  ylab("Temperature (C)") +
  plotTheme

#define plot layout
layout <- "
AA
AA
AA
BD
CD
CD"

#Merge plots
Fig1_ClimPlot = TempPlot + boxp + histo + Dens2D + plot_layout(design = layout)
Fig1_ClimPlot

##########
#Save PDFs
##########

pdf(paste(output.dir,"Fig1_Temperature deviation.pdf",sep="/"), width=9, height=6, useDingbats=FALSE)
Fig1_ClimPlot
dev.off()
```
</details>



<details><summary>Figure 2</summary>
```{r, warning=FALSE, message=FALSE, results="hide"}
#create panel labels
dat_text       = data.frame(
  label        = c("a","b","c","d","e","f","g","h","i"),
  phenology   = rep(c("Leaf_out","Senescence","GSL"),3),
  Species      = c(rep("Quercus",3), rep("Fagus",3), rep("Lonicera",3)),
  term         = "springwarming")
#Plot
Fig2_LMplot = ggplot(data = resultsLM.pheno, mapping = aes(x = term, y = estimate, 
                                                fill=phenology, alpha=term)) +
  geom_bar(position=position_dodge(), stat="identity", color="black") +
    geom_errorbar(aes(ymin=estimate-std.error, ymax=estimate+std.error),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9))+
  geom_hline(yintercept=0) +
  geom_text(data = dat_text, mapping = aes(x = Inf, y = Inf, 
                                           hjust = 1, vjust = 1,
                                           label = label,
                                           fontface = "bold"))+
  stat_summary(geom = 'text', label = resultsLM.pheno$sig, 
               fun.y = mean, vjust = 2) +
  scale_alpha_discrete(range = c(1, 0.7)) +
  xlab("Phenophase") +
  ylab("Effect size (days)") +
  scale_fill_manual(values = c("green3", "orange", "blue")) +
  facet_grid(Species~phenology) + 
  plotTheme4
Fig2_LMplot

#########
#Save PDF
#########

pdf(paste(output.dir,"Fig2_PhenologyLinearTreatmentModel.pdf",sep="/"), width=7, height=7, useDingbats=FALSE)
Fig2_LMplot
dev.off()
```
</details>



<details><summary>Figure 3</summary>
```{r, warning=FALSE, message=FALSE, results="hide"}
volume.data$Species = factor(volume.data$Species, levels=c("Quercus robur", "Fagus sylvatica", "Lonicera xylosteum"))
pd  = position_dodge(0.2) 
Fig3_LinePlot = ggplot(volume.data, aes(x=Year, y=Volume, colour=Treatment, group=Treatment)) + 
    stat_summary(fun = mean, geom="line", position=pd) +
    stat_summary(fun.data = "mean_se", size = 0.5, position=pd) +
    labs(x = "Year", y = "Stem volume (cm3)") +
    scale_color_manual(values=c("green3", "orange", "red3", "black")) +
    coord_cartesian(ylim = c(8, 152))+
    facet_grid(~Species) +
    plotTheme 
Fig3_LinePlot

#########
#Save PDF
#########

pdf(paste(output.dir,"Fig3_LinePlot.pdf",sep="/"), width=9, height=4, useDingbats=FALSE)
Fig3_LinePlot
dev.off()
```
</details>



<details><summary>Figure 4</summary>
```{r, warning=FALSE, message=FALSE, results="hide"}
#Ordering of factors
term           = c("springwarming", "autumnwarming", "springwarming:autumnwarming")
resultsLM.biomass$term = factor(resultsLM.biomass$term, levels=term, ordered=T)
#create panel labels
dat_text       = data.frame(
  label        = c("a","b","c","d","e","f","g","h","i","j","k","l"),
  organ        = rep(c("Total","Shoot","Root","RSR"),3),
  Species      = c(rep("Quercus",4), rep("Fagus",4), rep("Lonicera",4)),
  term         = "springwarming")
#Plot
LMplot = ggplot(data = resultsLM.biomass, mapping = aes(x = term, y = estimate, 
                                                fill=organ, alpha=term)) +
  geom_bar(position=position_dodge(), stat="identity", color="black") +
    geom_errorbar(aes(ymin=estimate-std.error, ymax=estimate+std.error),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9))+
  geom_hline(yintercept=0) +
  geom_text(data = dat_text, mapping = aes(x = -Inf, y = -Inf, 
                                           hjust = -0.3, vjust = -1,
                                           label = label,
                                           fontface = "bold"))+
  stat_summary(geom = 'text', label = resultsLM.biomass$sig, 
               fun.y = mean, vjust = 2) +
  scale_fill_manual(values=c("orange","yellow2","red3","grey70"))+
  scale_alpha_discrete(range = c(1, 0.7)) +
  xlab("Phenophase") +
  ylab("Effect size (%)") +
  facet_grid(Species~organ) + 
  plotTheme4
LMplot

#########
#Save PDF
#########

pdf(paste(output.dir,"Fig4_BiomassLinearTreatmentModel.pdf",sep="/"), width=8, height=7, useDingbats=FALSE)
LMplot
dev.off()
```
</details>



<details><summary>Figure 5</summary>
```{r, warning=FALSE, message=FALSE, results="hide"}
#Effect sizes
#############

#create panel labels
dat_text      = data.frame(
  label       = c("a","b","c","d","e","f","g","h","i","j","k","l"),
  organ       = rep(c("Total","Shoot","Root","RSR"),3),
  Species     = c(rep("Quercus",4), rep("Fagus",4), rep("Lonicera",4)),
  phenology   = "Leaf_out")

#Plot
LMplot = ggplot(data = resultsLM.pheno2, mapping = aes(x = phenology, y = estimate, 
                                                fill=organ, alpha=phenology)) +
  geom_bar(position=position_dodge(), stat="identity", color="black") +
    geom_errorbar(aes(ymin=estimate-std.error, ymax=estimate+std.error),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9))+
  geom_hline(yintercept=0) +
  geom_text(data = dat_text, mapping = aes(x = -Inf, y = -Inf, 
                                           hjust = -0.1, vjust = -1,
                                           label = label,
                                           fontface = "bold"))+
  stat_summary(geom = 'text', label = resultsLM.pheno2$sig, 
               fun.y = mean, vjust = 2) +
  scale_fill_manual(values=c("orange","yellow2","red3","grey70"))+
  scale_alpha_discrete(range = c(1, 0.7)) +
  xlab("Phenophase") +
  ylab("Effect size (%/day)") +
  facet_grid(Species~organ) + 
  plotTheme4
LMplot

#Biomass
########

#order
Organs      = c("Total","Shoot","Root","RSR") #order treatments
pheno.data3$organ = factor(pheno.data3$organ, levels=Organs, ordered=T)
#create panel labels
dat_text       = data.frame(
  label        = c("a","b","c","d","e","f","g","h","i"),
  phenology    = rep(c("Leaf_out","Senescence","GSL"),3),
  Species      = c(rep("Quercus",3), rep("Fagus",3), rep("Lonicera",3)),
  organ        = "Root")
BiomassPlot = ggplot(pheno.data3[pheno.data3$organ %in% c("Total","Shoot","Root"),], 
                     aes(x=phenologyAnomaly, y=relativeBiomass, colour=organ)) + 
    geom_hline(yintercept=0) +
    geom_vline(xintercept=0) +
    geom_smooth(method=lm, fullrange = T, alpha = 0.2) +
    labs(x = "Phenology anomaly (days)", y = "Biomass anomaly (%)") +
    scale_color_manual(values=c("orange","yellow2","red3"))+
    geom_text(data = dat_text, mapping = aes(x = -Inf, y = Inf, 
                                           hjust = -0.2, vjust = 1.3,
                                           label = label,
                                           fontface = "bold"), color="black")+
    geom_text(data    = resultsLM.pheno2[resultsLM.pheno2$organ=="Total",],
              mapping = aes(x = Inf, y = Inf, hjust = 1.5, vjust = 2.5, 
                            label = paste(round(estimate,1), " %/day", sig, sep="")), 
              size=3.5, color="orange")+
    geom_text(data    = resultsLM.pheno2[resultsLM.pheno2$organ=="Shoot",],
              mapping = aes(x = Inf, y = Inf, hjust = 1.5, vjust = 4.5, 
                            label = paste(round(estimate,1), " %/day", sig, sep="")), 
              size=3.5, color="yellow2")+
    geom_text(data    = resultsLM.pheno2[resultsLM.pheno2$organ=="Root",],
              mapping = aes(x = Inf, y = Inf, hjust = 1.5, vjust = 6.5, 
                            label = paste(round(estimate,1), " %/day", sig, sep="")), 
              size=3.5, color="red3")+
    facet_grid(Species~phenology, scales = "free") +
    plotTheme
BiomassPlot

#RSR
RSRplot = ggplot(pheno.data3[pheno.data3$organ %in% c("RSR"),], aes(x=phenologyAnomaly, y=biomass)) + 
    geom_point(size=0.2) +  
    geom_smooth(method=lm, fullrange = F) +
    labs(x = "Day", y = "Root:shoot ratio") +
    scale_color_viridis(option="viridis",discrete=TRUE) +
    geom_text(data    = resultsLM.pheno2[resultsLM.pheno2$organ=="RSR",],
              mapping = aes(x = Inf, y = Inf, hjust = 1.5, vjust = 2.5, 
                            label = paste("Slope = ", round(estimate,3), " day-1", sig, sep="")), 
              size=3.5, color="black")+
    facet_grid(Species~phenology, scales = "free") +
    plotTheme
RSRplot


#########
#Save PDF
#########

pdf(paste(output.dir,"Fig5_PhenologyBiomass.pdf",sep="/"), width=8, height=7, useDingbats=FALSE)
BiomassPlot
RSRplot
dev.off()
```
</details>



<details><summary>Figure S1</summary>
```{r, warning=FALSE, message=FALSE, results="hide"}
#create panel labels
dat_text <- data.frame(
  label       = c("a","b","c","d","e","f","g","h","i"),
  phenology   = rep(c("Leaf_out","Senescence","GSL"),3),
  Species     = c(rep("Quercus",3), rep("Fagus",3), rep("Lonicera",3)),
  Treatment   = "spring")
#Plot
FigS1_PhenologyChangePlot = ggplot(data = pheno.data2, mapping = aes(x = Treatment, y = phenologyChange, 
                                                         fill=phenology, alpha=Treatment)) +
  stat_summary(fun.y = mean, 
                 geom = "bar", color="black") + 
  stat_summary(fun.data = mean_se,  
                 geom = "errorbar", width=0.2) + 
  geom_hline(yintercept=0) +
  geom_text(data = dat_text, mapping = aes(x = -Inf, y = Inf, 
                                           hjust = -0.2, vjust = 1.5,
                                           label = label,
                                           fontface = "bold"))+
  stat_summary(geom = 'text', label = resultsTT$significance, 
               fun.y = mean, vjust = 2) +
  scale_alpha_discrete(range = c(1, 0.7)) +
  xlab("Treatment") +
  ylab("Phenological change (days)") +
  scale_fill_manual(values = c("green3", "orange", "blue")) +
  facet_grid(Species~phenology) + 
  plotTheme4
FigS1_PhenologyChangePlot

#########
#Save PDF
#########

pdf(paste(output.dir,"FigS1_PhenologyChange.pdf",sep="/"), width=7, height=7, useDingbats=FALSE)
FigS1_PhenologyChangePlot
dev.off()
```
</details>



<details><summary>Figure S2</summary>
```{r, warning=FALSE, message=FALSE, results="hide"}
#Plot
#####

#create panel labels
dat_text <- data.frame(
  label = c("a","b","c","d","e","f","g","h","i","j","k","l"),
  organ   = rep(c("Total","Shoot","Root","RSR"),3),
  Species     = c(rep("Quercus",4), rep("Fagus",4), rep("Lonicera",4)),
  Treatment   = "spring")
#Plot
BiomassChangePlot = ggplot(data = biomass.data1, mapping = aes(x = Treatment, y = relativeBiomass, 
                                                       fill=organ, alpha=Treatment)) +
  stat_summary(fun.y = mean, 
                 geom = "bar", color="black") + 
  stat_summary(fun.data = mean_se,  
                 geom = "errorbar", width=0.2) + 
  geom_hline(yintercept=0) +
  geom_text(data = dat_text, mapping = aes(x = -Inf, y = -Inf, 
                                           hjust = -0.3, vjust = -1,
                                           label = label,
                                           fontface = "bold"))+
  stat_summary(geom = 'text', label = resultsTTbiomass$sig, 
               fun.y = max, vjust = 2.9) +
  scale_alpha_discrete(range = c(1, 0.7)) +
  coord_cartesian(ylim = c(-60, 60))+
  xlab("Treatment") +
  ylab("Biomass change (%)") +
  scale_fill_manual(values=c("orange","yellow2","red3","grey70"))+
  facet_grid(Species~organ) + 
  plotTheme4
BiomassChangePlot

#########
#Save PDF
#########

pdf(paste(output.dir,"FigS2_BiomassChange.pdf",sep="/"), width=8, height=7, useDingbats=FALSE)
BiomassChangePlot
dev.off()
```
</details>

<br>

## 6. Reproducibility

<details><summary>Reproducibility receipt</summary>

```{r}
## datetime
Sys.time()

## repository
#git2r::repository()

## sessioninfo
sessionInfo()
```

</details>


